---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bplsr

<!-- badges: start -->
<!-- badges: end -->

The `bplsr` package implements the Bayesian partial least squares regression model. It is a Bayesian factor model which emulates the partial least squares (PLS) method. See Urbas et al. (2024) for details.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("SzymonUrbas/bplsr")
```

## Example

The following example illustrates how to carry out multivariate regression using BPLS on mid-infrared spectral data of milk samples:

```{r example}
library(bplsr)


X = milk_MIR$xMIR
Y = milk_MIR$yTraits[, c('Casein_content','Fat_content')]

set.seed(1)
# fit model to 75% of data and predict on remaining 25%
idx = sample(seq(nrow(X)),floor(nrow(X)*0.75),replace = FALSE)

Xtrain = X[idx,];Ytrain = Y[idx,]
Xtest = X[-idx,];Ytest = Y[-idx,]

# fit the model (MCMC takes time)
bplsr_Fit = bplsr(Xtrain,Ytrain)

# generate predictions
bplsr_pred = bplsr.predict(model = bplsr_Fit, newdata = Xtest)

# point predictions
head(bplsr_pred$Ytest)

# lower and upper limits of prediction interval
head(bplsr_pred$Ytest_PI)

# plot of predictive posterior distribution for single test sample
hist(bplsr_pred$Ytest_dist[1,'Casein_content',], freq = F,
     main = 'Posterior predictive density', xlab = 'Casein_content')

```


### References
Urbas, S., Lovera, P., Daly, R., O'Riordan, A., Berry, D., and Gormley, I. C. (2024). "Predicting milk traits from spectral data using Bayesian probabilistic partial least squares regression." _The Annals of Applied Statistics_, 18(4): 3486-3506 [doi:10.1214/24-AOAS1947](https://doi.org/10.1214/24-AOAS1947)